Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-26 paper/ijcai16.tex
Introduction

%The problem is that of learning or assessing the preferences of the user (or decision maker).
Preferences \cite{Peintner2008} play an important role in artificial intelligence,
including fields spanning from recommender systems to automatic planning, 
from non monotonic reasoning to computational social choice and algorithmic decision theory.
The elicitation of learning of preferences is therefore a crucial task, .
The goal in preference elicitation is that of gaining enough information about the user's preferences (typically represented by an utility function $u$, unknown to the system) in order to recommend a decision.

While preference elicitation has been often cast as that of adapting using ad-hoc rules \cite{},
recently a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas-mmr:uai07,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.
One option is to be Bayesian and to maintain a probability distribution over the possible realizations of the utility's parameters \cite{chajewska2000,boutilier2002,guo2010real,viappiani2010optimal}.
Alternatively, when distribution information about the parameters is not available, one can reason about all feasible utility functions consistent with the currently known information about the user; from the user's response we are able to infer constraints on the parameters implicated in the utility model \cite{Wang2003,boutilier2006,braziunasmmr:uai07,viappiani2009}.
This latter approach has also the advantage of avoiding the computationally demanding task of maintaining distribution information and performing Bayesian updates.

% constructive preference elicitation
In this paper we take a {\em constructive} view on preference
elicitation, enlarging its scope from the selection of items among a
set of candidates to the synthesis of entirely novel
instances. Instances are represented as combinations of basic elements
(e.g. the components of a laptop) subject to a set of constraints
(e.g. the laptop model determines the set of available CPUs). A
utility function is learned over the feature representation of an
instance, as customary in many preference elicitation approaches. The
recommendation is then made by solving a constrained optimization
problem in the space of feasible instances, guided by the learned
utility.

%\andrea{other constructive preference elicitation approaches?}

% setwise max-margin formulation to deal with uncertainty in user utility
Learning a user utility function requires setting a preference
elicitation strategy and dealing with uncertain and possibly
inconsistent user feedback. Bayesian preference elicitation approaches
deal with both problems by building a probability distribution on
candidate functions and asking queries maximizing informativeness
measures such as expected value of information (EVOI)~\cite{}. These
approaches are however computationally expensive and cannot scale to
fully constructive scenarios, as shown in our experimental results.
We take a space decomposition perspective and jointly learn a set of
weight vectors, each representing a candidate utility function,
maximizing diversity between the vectors and consistency with the
available feedback. These two conflicting objectives tend to generate
equally plausible alternative hypotheses for the unknown
utility. 
Our approach to preference elicitation works by combining weight vector
learning with instance generation, so that each iteration of the
algorithm produces two outcomes: a set of weight vectors and a set of
instances, each maximizing its score according to one of the weight
vectors.

@@

The paper is structured as follows: after discussing related works in next Section,
we present some brackground material in Section\ref{sec:background} 
and then proceed to our setwise maxmargin method in Section\ref{sec:formulationm}; 
we discuss experimental results in Section\ref{sec:experiments}
and conclude with final remarks (Section\ref{sec:conclusions}).

\section{Related Work}

Preference elicitation for the customization of user interfaces \cite{gajos2005}.

WRITEME

\section{Background}
\ref{sec:background}
%\section{Preference Elicitation}

@@

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=16em]{figures/setmargin}
    \end{center}
    \caption{\label{fig:setmargin} Intuition.}
\end{figure}

@@

\paragraph{Set-wise max-margin.} The full {\sc SetMargin} algorithm
follows the usual preference elicitation loop. Starting from an
initially empty set of user responses $\calD$, it repeatedly solves
the MILP problem above using the set $\calD$ to enforce ranking
constraints on the weight vectors $\{\vw^i\}$. The generated
configurations $\{\vx^i\}$, which are chosen to be as good as possible
with respect to the estimated user preferences, and as diverse as
possible, are then employed to formulate a set of user queries. The
new replies are added to $\calD$ and the whole procedure is
repeated. The algorithm terminates after a fixed number of iterations,
or once another suitable condition is met. \andrea{@Paolo: what termination
condition for pref elicitation?}  The procedure is sketched in
Algorithm~\ref{alg:setmargin}. Note that at the end of the preference
elicitation procedure, a final recommendation is made by solving the
MILP problem for $k=1$.

@@

We implemented the {\sc SetMargin} algorithm using Python, leveraging Gurobi
6.5.0 for solving the core MILP problem. Both the {\sc SetMargin} source code
and the full experimental setup are available at \stefano{add URL}.

@@

\section{Conclusion}

WRITEME

>>>>

Preferences \cite{Peintner2008} play an important role in artificial intelligence,
including fields spanning from recommender systems to automatic planning, 
from non monotonic reasoning to computational social choice and algorithmic decision theory.
The task of eliciting or learning preferences is therefore a crucial; typically only limited information about the user's preferences will be available and the cost of obtaining additional preference information will be usually high (from a cognitive and/or computational point of view).
The goal in preference elicitation is that of gaining enough information about the user's preferences (typically represented by an utility function $u$, unknown to the system) in order to recommend a decision or a course of action.

The assessment of preferences have received considerable attention in the literature, with early works confined to the operation research and management literature mostly \cite{}.
In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective; within the machine learning field the term preference learning has been catching up.
While preferences are often assessed using some ad-hoc rules \cite{},
recently a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas-mmr:uai07,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.

%One option is to be Bayesian and to maintain a probability distribution over the possible realizations of the utility's parameters \cite{chajewska2000,boutilier2002,guo2010real,viappiani2010optimal}.
%Alternatively, when distribution information about the parameters is not available, one can reason about all feasible utility functions consistent with the currently known information about the user; from the user's response we are able to infer constraints on the parameters implicated in the utility model \cite{Wang2003,boutilier2006,braziunasmmr:uai07,viappiani2009}.
%This latter approach has also the advantage of avoiding the computationally demanding task of maintaining distribution information and performing Bayesian updates.

% constructive preference elicitation
While most works assume that items or decisions are available in a (possibly large)
dataset, in this paper we propose an adaptive elicitation framework that takes a {\em constructive} view on preference
elicitation, enlarging its scope from the selection of items among a
set of candidates to the synthesis of entirely novelinstances. 
Instances are solutions to a given optimization problem; they are represented as combinations of basic elements
(e.g. the components of a laptop) subject to a set of constraints
(e.g. the laptop model determines the set of available CPUs). 
A utility function is learned over the feature representation of an
instance, as customary in many preference elicitation approaches. 
The recommendation is then made by solving a constrained optimization
 problem in the space of feasible instances, guided by the learned
utility.

Our approach is related to \cite{boutilier2006}, where minimax regret is used both
as a robust recommendation criterion and as a technique to drive elicitation in configuration problems.
The main limitation of the minimax regret approach is the lack of tolerance with respect to user inconsistency (the user might state that item x is preferred to y, even though the latter has higher utility).

% setwise max-margin formulation to deal with uncertainty in user utility
Indeed, learning a user utility function requires setting a preference
elicitation strategy and dealing with uncertain and possibly
inconsistent user feedback. 
Bayesian preference elicitation approaches
deal with both problems by building a probability distribution on
candidate functions (endowed with a response or error model to be used for inference) and asking queries maximizing informativeness
measures such as expected value of information (EVOI)~\cite{chajewska2000,viappiani2010optimal}.

These
approaches are however computationally expensive and cannot scale to
fully constructive scenarios, as shown in our experimental results.
We take a space decomposition perspective and jointly learn a set of
weight vectors, each representing a candidate utility function,
maximizing diversity between the vectors and consistency with the
available feedback. These two conflicting objectives tend to generate
equally plausible alternative hypotheses for the unknown
utility. 
Our approach to preference elicitation works by combining weight vector
learning with instance generation, so that each iteration of the
algorithm produces two outcomes: a set of weight vectors and a set of
instances, each maximizing its score according to one of the weight
vectors.

@@

The paper is structured as follows: after discussing related works in next Section,
we present some brackground material in Section \ref{sec:background} 
and then proceed to our setwise maxmargin method in Section \ref{sec:formulation}; 
we discuss experimental results in Section \ref{sec:experiments}
and conclude with final remarks (Section \ref{sec:conclusions}).

%\section{Related Work}

%Preference elicitation for the customization of user interfaces \cite{gajos2005}.

%WRITEME

\section{Background}
\label{sec:background}
%\section{Preference Elicitation}

@@

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=16em]{figures/setmargin}
    \end{center}
    \caption{\label{fig:setmargin} The intuition behind setwise max-margin; the red points are utilty vectors $\vw^{i}$ $\vw^{j}$, the red line corresponds to the constraint $\langle \vw, \vx^{1} - \vx^{2} \rangle \ge 0$.}
\end{figure}

@@

\paragraph{Set-wise max-margin.} The full {\sc SetMargin} algorithm
follows the usual preference elicitation loop. Starting from an
initially empty set of user responses $\calD$, it repeatedly solves
the MILP problem above using the set $\calD$ to enforce ranking
constraints on the weight vectors $\{\vw^i\}$. The generated
configurations $\{\vx^i\}$, which are chosen to be as good as possible
with respect to the estimated user preferences, and as diverse as
possible, are then employed to formulate a set of user queries. The
new replies are added to $\calD$ and the whole procedure is
repeated. The algorithm terminates after a fixed number of iterations,
or once another suitable condition is met. \andrea{@Paolo: what termination
condition for pref elicitation?}  
Termination might be left to the user to decide (see for example \ref{} for discussion on user-involved recommender systems); the system can however propose
to terminate when the difference between the utility weights is small enough.

The procedure is sketched in
Algorithm~\ref{alg:setmargin}. Note that at the end of the preference
elicitation procedure, a final recommendation is made by solving the
MILP problem for $k=1$.

@@

We implemented the {\sc SetMargin} algorithm using Python, leveraging Gurobi
6.5.0 for solving the core MILP problem. Both the {\sc SetMargin} source code
and the full experimental setup are available at {\tt https://db.tt/oqY62odW}.

 \stefano{add URL}.

@@

\section{Conclusion}
\label{sec:conclusions}

This paper presented a maxmargin approach to generating a set of diverse examples.
Our approach, that can be seen as an extension of maxmargin learning to sets, is effective in the generation of a diverse set of items that can be used to ask informative queries to the user.
The main advantages of our elicitation method are 1) its aptitude for eliciting preferences and providing recommendations in large configuration problems  2) robustness with respect to erroneous preference input, and 3) its ability to support sparse utility functions.
In a number of experiments we showed how our method compare with state-of-the-art approaches.

\paolo{any idea about possible future works?}
% FUTURE WORKS
